---
title: "Project_II"
author: "Brenda Maldonado"
date: "9/28/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###### Exercise 2: Analysis of Variance
###### For this problem use the bupa.csv data set. Check UCI Machine Learning Repository for more information (http://archive.ics.uci.edu/ml/datasets/Liver+Disorders). The mean corpuscular volume and alkaline phosphatase are blood tests thought to be sensitive to liver disorder related to excessive alcohol consumption. We assume that normality and independence assumptions are valid.

```{r}
bupa = read.csv('bupa.csv', header=TRUE)
head(bupa,2)
```

###### a) Perform a one-way ANOVA for mcv as a function of drinkgroup. Comment on significance of the drinkgroup, the amount of variation described by the model, and whether or not the equal variance assumption can be trusted.

```{r}
aov(mcv~drinkgroup,bupa)
```
###### Due to the residual standard error being 4.25495, this model may not be a good fit and the estimated effect may be unbalanced.

######b) Perform a one-way ANOVA for alkphos as a function of drinkgroup. Comment on statistical significance of the drinkgroup, the amount of variation described by the model, and whether or not the equal variance assumption can be trusted.

```{r}
aov(alkphos~drinkgroup,bupa)
```
###### Due to the residual standard error being 18.34899, this model is a worse fit than mcv model since the residual standard error is higher. The estimated effects may be unbalanced. 
